---
title: "Does density affect infection load?"
output: html_notebook
---

```{r}
library(data.table)
library(rstan)
library(ggplot2)
library(stringr)
library(brms)
```

## Analysis III: Does host density affect infection load?

```{r}
anal_dat = fread("../data/candidate_datasets/analysisIII_loadanddensity.csv")
```

## Run a SEM analysis on the data

The hypothesized model is

$$
\begin{aligned}
  Host_{t} &= \beta_{11} Bd_{t - 1} + \epsilon_1 \\
  Bd_{t} &= \beta_{21} Bd_{t - 1} + \beta_{22} Host_t + \epsilon_2
\end{aligned}
$$

The full covariance matrix has 6 unique parameters so that would be the fully specified model. 


```{r}
# Location == 1 refers to yosemite populations
yose_dat = anal_dat[location == 1]

# Regression model with log10 frog abundance
sm1_abund = lm(scale(mean_bd_present) ~ scale(mean_bd_past) + scale(log10(frog_abund_present)), data=yose_dat)
bayes1_abund = brm(scale(mean_bd_present) ~ scale(mean_bd_past) + scale(log10(frog_abund_present)), data=yose_dat)
sm_abund_no_trans = lm(mean_bd_present ~ mean_bd_past + log10(frog_abund_present), data=yose_dat)
# plot(sm1_abund)
summary(sm1_abund)
confint(sm1_abund)
bayes1_abund

# Regression model with log10 frog density
sm1_density = lm(scale(mean_bd_present) ~ scale(mean_bd_past) + scale(log10(density)), data=yose_dat)
bayes1_density = brm(scale(mean_bd_present) ~ scale(mean_bd_past) + scale(log10(density)), data=yose_dat)
sm_density_no_trans = lm(mean_bd_present ~ mean_bd_past + log10(density), data=yose_dat)
# plot(sm1_density)
summary(sm1_density)
confint(sm1_density)

# Regression model with past mean Bd affecting abundance
sm2_abund = lm(scale(log10(frog_abund_present)) ~ scale(mean_bd_past), data=yose_dat)
summary(sm2_abund)
confint(sm2_abund)

# Regression model with present mean Bd affecting abundance
sm2_density = lm(scale(log10(density)) ~ scale(mean_bd_past), data=yose_dat)
summary(sm2_density)
confint(sm2_density)
```

Plot partial regression plots to display results

```{r}
library(patchwork)

# Plot of host abundance

residx = resid(lm(scale(log10(frog_abund_present)) ~ scale(mean_bd_past), data=yose_dat))
residy = resid(lm(scale(mean_bd_present) ~ scale(mean_bd_past), data=yose_dat))

p1 = ggplot() + geom_point(aes(x=residx, y=residy)) + stat_smooth(aes(x=residx, y=residy), method="lm", color='red') +
                theme_classic() + xlab("Host abundance residuals") + ylab("Bd intensity residuals")

residx_dens = resid(lm(scale(log10(density)) ~ scale(mean_bd_past), data=yose_dat))
residy_dens = resid(lm(scale(mean_bd_present) ~ scale(mean_bd_past), data=yose_dat))

p2 = ggplot() + geom_point(aes(x=residx_dens, y=residy_dens)) + stat_smooth(aes(x=residx_dens, y=residy_dens), method="lm", color='red') +
                theme_classic() + xlab("Host density residuals") + ylab("Bd intensity residuals")

fplot = p1 + p2 + plot_annotation(tag_levels = "A", tag_suffix = ".")
fplot
ggsave("../results/within_host_plot.pdf", width=7, height = 3.5)

```

## Translocation analysis

Are load trajectories significantly different on a single host experiencing exponential pathogen growth compared to hosts in a high density population experiencing pathogen growth?

```{r}
library(ggplot2)
library(data.table)
library(rstan)
```

```{r}
trans_data = fread("../data/archival/humphreys_translocation_2010.csv")
trans_data[, visit_date:=as.POSIXct(visit_date, format="%m/%d/%y")]

full_data = fread("../data/formatted/combined_Bd_data_for_analysis_allVES.csv")
full_data[, visit_date:=as.POSIXct(visit_date, format="%m/%d/%y")]
```

```{r}
site_dat = full_data[site_id == 50183 & year == 2010][, .(mean_bd=mean(log10(bd_load + 1))), by=.(visit_date)]
site_dat
```

```{r}

ggplot(data=trans_data[!is.na(pit_tag_ref)]) + 
        geom_point(aes(x=visit_date, y=log10(bd_load + 1), group=pit_tag_ref, color=as.factor(site_id))) + 
        geom_line(aes(x=visit_date, y=log10(bd_load + 1), group=pit_tag_ref, color=as.factor(site_id))) + theme_classic() +
        geom_point(data=site_dat, aes(x=visit_date, y=mean_bd), shape=4) + 
        geom_line(data=site_dat, aes(x=visit_date, y=mean_bd)) +
        geom_point(data=full_data[site_id == 50183 & year == 2010], aes(x=visit_date, y=log10(bd_load + 1)))
  
```
We want to test whether the maximum load for individuals in 50219, 50194, and 50170 differs from the maximum load in 50813

```{r}

# Subset and combine data
swab_data = full_data[site_id == 50183 & year == 2010 & visit_date == as.POSIXct("2010-09-07", tz="UTC")]
indiv_data = trans_data[site_id %in% c(50219, 50194, 50170) & visit_date > as.POSIXct("2010-09-07", tz="UTC")]
comb_data = rbind(swab_data[ ,.(site_id, bd_load, density="high", source="donor")], 
                  indiv_data[, .(site_id, bd_load, density="low", source="recipient")])
comb_data$site_id = factor(comb_data$site_id, levels=c(50183, 50170, 50194, 50219))

# remove outliers to test
#comb_data = comb_data[log10(bd_load) > 5]
```

Fit a simple GLS model

```{r}
library(nlme)

fit_null = gls(log10(bd_load) ~ 1, data=comb_data, method="ML")
fit1 = gls(log10(bd_load) ~ density, data=comb_data, method="ML")

varfunc = varIdent(form=~1|as.factor(site_id))
fit2 = gls(log10(bd_load) ~ density, 
           weights=varfunc, data=comb_data, method="ML")

fit3 = gls(log10(bd_load) ~ 1, 
           weights=varfunc, data=comb_data, method="ML")

fit4 =  gls(log10(bd_load) ~ as.factor(site_id), 
           weights=varfunc, data=comb_data, method="ML")

summary(fit2)
AIC(fit_null, fit1, fit2, fit3, fit4)

summary(fit4)
unique(comb_data$site_id)
```

Fit a Bayesian version with Stan

```{r}

mod1 = "
data{

  int N;
  int P;
  int D;
  real y[N];
  matrix[N, P] X;
  matrix[N, D] S;
} 
parameters{
  real beta0;
  vector[P] beta;
  vector<lower=0>[D] sigma;
} 
model{
  y ~ normal(beta0 + X*beta, S*sigma);
} generated quantities{
 real log_lik[N];
 
 for(i in 1:N){
  log_lik[i] = normal_lpdf(y[i] | beta0 + row(X, i)*beta, row(S, i)*sigma);
 }
}
"

mod_greater = "
data{

  int N;
  int P;
  int D;
  real y[N];
  matrix[N, P] X;
  matrix[N, D] S;
} 
parameters{
  real beta0;
  vector<lower=0>[P] beta;
  vector<lower=0>[D] sigma;
} 
model{
  y ~ normal(beta0 + X*beta, S*sigma);
} generated quantities{
 real log_lik[N];
 
 for(i in 1:N){
  log_lik[i] = normal_lpdf(y[i] | beta0 + row(X, i)*beta, row(S, i)*sigma);
 }
}
"

mod_less = "
data{

  int N;
  int P;
  int D;
  real y[N];
  matrix[N, P] X;
  matrix[N, D] S;
} 
parameters{
  real beta0;
  vector<lower=0>[P] beta;
  vector<lower=0>[D] sigma;
} 
model{
  y ~ normal(beta0 - X*beta, S*sigma);
} generated quantities{
 real log_lik[N];
 
 for(i in 1:N){
  log_lik[i] = normal_lpdf(y[i] | beta0 - row(X, i)*beta, row(S, i)*sigma);
 }
}
"
```

```{r}
# Build data for stan
tfit = lm(log10(bd_load) ~ as.factor(site_id), data=comb_data)
X = model.matrix(tfit)
X = X[, 2:ncol(X)]
S = as.matrix(mltools::one_hot(as.data.table(comb_data$site_id)))
```

```{r}
# Compile models
smod = stan_model(model_code=mod1)
smod_greater = stan_model(model_code=mod_greater)
smod_less = stan_model(model_code=mod_less)
```

```{r}
standata = list(N=nrow(X), P=ncol(X),
                D=ncol(S), 
                y=log10(comb_data$bd_load),
                X=X,
                S=S)
# Fit models
bfit1 = rstan::sampling(smod, data=standata, cores=4) # Means free to vary
bfit_greater = rstan::sampling(smod_greater, data=standata, cores=4) # Recipient means greater than donor
bfit_less = rstan::sampling(smod_less, data=standata, cores=4) # Recipient means less than donor
```

```{r}
bfit1
```

Model comparison with LOO

```{r}
library(loo)

loglik1 = extract_log_lik(bfit1, merge_chains = FALSE)
loglik2 = extract_log_lik(bfit_greater, merge_chains = FALSE)
loglik3 = extract_log_lik(bfit_less, merge_chains = FALSE)

reff1 = relative_eff(exp(loglik1)) 
reff2 = relative_eff(exp(loglik2))
reff3 = relative_eff(exp(loglik3))

loo1 = loo(loglik1, r_eff=reff1)
loo2 = loo(loglik2, r_eff=reff2)
loo3 = loo(loglik3, r_eff=reff3)

comp = loo_compare(loo1, loo2, loo3)

# View results
comp
loo1
loo2
loo3

```
Plot all translocation results together

```{r}
library(patchwork)

colors = c('#1f78b4','#33a02c')
colors2 = c('#33a02c', '#1f78b4', '#33a02c', '#33a02c')
samp_size = comb_data[, .(n_samps=paste0("n=", .N)), by=.(site_id)]
samp_size$sig = c("", "**", "ns" ,"**")

# Extract lake perimeters to calculate density
lake_dat = fread("../data/archival/sites_nov2020.csv")
lake_dat_trun = lake_dat[id %in% unique(comb_data$site_id)][ , .(area=area, perimeter=2*sqrt(area / pi)*pi, site_id=id)]
lake_dat_trun


tplot = ggplot(comb_data) + geom_boxplot(aes(x=site_id, y=log10(bd_load), fill=source)) + 
                            geom_text(data=samp_size, aes(x=site_id, y=3, label=n_samps)) +
                            geom_text(data=samp_size, aes(x=site_id, y=7.5, label=sig), size=6) +
                            theme_classic() + ylab("Bd infection intensity\nlog10(Bd load)") + xlab("Population ID") +
                            scale_fill_manual(values=colors) +
                            guides(fill=guide_legend(title=element_blank())) + 
                            theme(legend.position = "none")

# Assign lake id as the final lake id
map = c("50183"="donor", "50170"="translocated", "50194"="translocated", "50219"="translocated")
ind_map = trans_data[!is.na(pit_tag_ref) & site_id != 50183][, .(id=unique(site_id)), by=.(pit_tag_ref)]
trans_data_up = merge(trans_data, ind_map, key="pit_tag_ref", all.x=T)
trans_data_up$donor = plyr::revalue(as.character(trans_data_up$site_id), map)
pit_mean = trans_data_up[!is.na(pit_tag_ref) & !is.na(id)][, .(mean_bd=mean(log10(bd_load + 1))), by=.(visit_date, id)]
pit_mean$donor = plyr::revalue(as.character(pit_mean$id), map)

tplot2 = ggplot(data=trans_data_up[!is.na(pit_tag_ref) & !is.na(id)]) + 
        geom_point(aes(x=visit_date, y=log10(bd_load + 1), group=pit_tag_ref, color=as.factor(id)),
                   alpha=0.15) + 
        geom_line(aes(x=visit_date, y=log10(bd_load + 1), group=pit_tag_ref, color=as.factor(id)), alpha=0.15) +
        geom_point(data=pit_mean, aes(x=visit_date, y=mean_bd, color=as.factor(id)), shape=18, size=6) +
        geom_line(data=pit_mean, aes(x=visit_date, y=mean_bd, color=as.factor(id)), size=1) +
        geom_point(data=site_dat, aes(x=visit_date, y=mean_bd, color="50183"), shape=18, size=6) + 
        geom_line(data=site_dat, aes(x=visit_date, y=mean_bd, color="50183"), size=1) +
        geom_point(data=full_data[site_id == 50183 & year == 2010],
                   aes(x=visit_date, y=log10(bd_load + 1), color="50183"), alpha=0.15) +
        scale_color_manual(values=colors2) +
        xlab("Visit date in 2010") + ylab("Bd infection intensity\nlog10(Bd load + 1)") +
        guides(color=guide_legend(title=element_blank())) + theme_classic() + 
        theme(legend.position = "none", 
              legend.background = element_rect(fill="transparent"),
              axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Make plot of maximum adult abundance based on VES
count_data = fread("../data/archival/ramu_counts_nov2020.csv")
count_data[, visit_date:=as.POSIXct(visit_date, format="%m/%d/%y")][, year:=year(visit_date)]
site_ids = unique(comb_data$site_id)
max_counts = count_data[site_id %in% site_ids & year == 2010][, .(max_count=max(stage_count)), by=.(site_id, year, visual_life_stage)]

max_counts$site_id = factor(max_counts$site_id, levels=c(50183, 50170, 50194, 50219))
max_counts$source = c("translocated", "donor", "donor", "donor", "translocated", "translocated")
lake_dat_trun$site_id = factor(lake_dat_trun$site_id, levels=c(50183, 50170, 50194, 50219))
max_counts = merge(max_counts, lake_dat_trun, by="site_id", all.x=TRUE)

diff = 0.17
labels = data.frame(x=c(1 - diff, 1, 1 + diff, 2 - diff, 3 - diff, 4 - diff), 
                    y=rep(10, 6), 
                    label=c("Adult", "Subadult", "Tadpole","Adult", "Adult", "Adult"))
tplot3 = ggplot(max_counts) + 
            geom_bar(aes(x=as.factor(site_id), y=max_count, fill=source, group=visual_life_stage), 
                     stat='identity', position=position_dodge(preserve = "single"), width=0.5, color='black') +
            scale_y_log10() +
            theme_classic() + ylab("Maximum abundance in 2010") + xlab("Population ID") +
            scale_fill_manual(values=colors) +
            guides(fill=guide_legend(title=element_blank()), ) + 
            theme(legend.position = c(0.75, .9))
tplot3 = tplot3 + geom_text(data=labels, aes(x=x, y=y, label=label), angle=90, size=2.5)



layout = "
AA
BC
"
all_plot = tplot2 + tplot + tplot3 + plot_layout(design=layout) + plot_annotation(tag_levels = "A", tag_suffix = ".")
all_plot
ggsave("../results/combine_translocation_plot.pdf", plot=all_plot, width=8, height=8)
```
